Translation Elicitation

### 2021-07-01 Rmd/report.Rmd
Rename files, include job, add index.html

---
title: "Translation Elicitation"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

# import data
trials <- readRDS(here("Data", "trials.rds"))
participants <- readRDS(here("Data", "participants.rds"))
responses <- readRDS(here("Data", "responses.rds"))
jtrace <- readRDS(here("Data", "jtrace.rds"))

@@

```{r participants_trials, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

Participants in the ENG-SPA listened to `r nrow(trials[trials$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(trials[trials$group=="ENG-CAT",])` trials. Five trials in each condition were practice trials and were excluded from analyses.

```{r trials_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
trials_coded <- trials %>% 
  mutate(
    test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"),
    zipf = log10(frequency)+3,
    pthn_log = log(pthn),
    lv_log = log(lv)
  )
```

@@

```{r trials_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials_coded %>% 
  ggplot(aes(zipf, fill = test_language)) +
  geom_histogram(bins = 25) +
  labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme_custom() +
  theme(
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

@@

```{r trials_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials_coded %>% 
  count(test_language, pthn) %>% 
  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
  geom_col(position = position_dodge()) +
  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

@@

```{r trials_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials_coded %>% 
  ggplot(aes(group, lv, fill = group, colour = group)) +
  geom_violin() +
  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
  labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  ) +

  trials_coded %>% 
  select(trial_id, group, ends_with("_ratio")) %>% 
  pivot_longer(ends_with("_ratio"), names_to = "type", values_to = "ratio") %>% 
  mutate(type = str_to_sentence(str_remove(type, "_ratio"))) %>% 
  ggplot(aes(group, ratio, fill = group, colour = group)) +
  facet_wrap(~type, ncol = 1) +
  geom_violin() +
  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
  labs(x = "Grouop", y = "Similarity", fill = "Test language") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  ) +

@@

```{r trials_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials_coded %>% 
  select(trial_id, group, onset) %>% 
  ggplot(aes(group, fill = onset, colour = onset)) +
  geom_bar(position = position_dodge(width = 0.9)) +
  labs(x = "Group", y = "# trials", fill = "Phonological onset", colour = "Phonological onset") +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

@@

```{r trials_similarity_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials_coded %>% 
  select(trial_id, group, lv, onset) %>% 
  ggplot(aes(lv, fill = onset, colour = onset)) +
  facet_wrap(~group, ncol = 2) +
  geom_histogram() +
  labs(x = "Phonological Leveshtein distance\n(corrected for string length)", y = "# trials",
       fill = "Phonological onset", colour = "Phonological onset") +
  theme_custom() +
  theme(
    legend.position = c(0.75, 0.25),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

@@

m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
  mutate_at(vars(lv), function(x) paste0(x, " SD"))

ggplot(m, aes(pthn, .value, colour = lv, fill = lv)) +
  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(x = "PTHN", y = "P(Correct)", colour = "Phon. similarity", fill = "Phon. similarity") +
  scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid = element_line(colour = "grey", linetype = "dotted")
  ) 
ggsave(here("Figures", "responses_pthn.png"))

@@

m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
  mutate_at(vars(pthn), function(x) paste0(x, " SD"))

ggplot(m, aes(lv, .value, colour = pthn, fill = pthn)) +
  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(x = "Phon. similarity", y = "P(Correct)", colour = "PTHN", fill = "PTHN") +
  scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid = element_line(colour = "grey", linetype = "dotted")
  ) +
  ggsave(here("Figures", "responses_lv.png"))

@@

```{r jtrace_fixed_table, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
summary(fit_jtrace$fit)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_number(2:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="(Intercept)")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "consonant_ratio", "vowel_ratio", "onset1", "overlap_stress1") | str_detect(term, "activation"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  )
```

```{r jtrace_fixed_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%"}
fit_jtrace$fixed %>%
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
  ggplot(aes(.value, fill = .variable)) +
  facet_wrap(~.variable, scales = "free") +
  geom_vline(xintercept = 0, colour = "red") +
  stat_halfeye() +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
    axis.text = element_text(size = 7)
  )
```

### Random coefficients

```{r jtrace_post_re, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
re <- ranef(fit_jtrace$fit)$`group:input`
re <- rbind(Intercept = re[,,1], activation = re[,,2]) %>% 
  as.data.frame() %>% 
  rownames_to_column("term") %>% 
  mutate(
    term = str_replace(str_remove(term, ".1"), "\\.", "\\-"),
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    Q2.5 = ifelse(term=="Intercept", inv_logit_scaled(Q2.5), Q2.5/4),
    Q97.5 = ifelse(term=="Intercept", inv_logit_scaled(Q97.5), Q97.5/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  separate(term, c("group", "input"), sep = "_")
re$.param <- rep(c("Intercept", "activation"), each = 9)

ggplot(re, aes(Estimate, input, colour = input)) +
  facet_wrap(~.param) +
  geom_vline(xintercept = 0, colour = "red") +
  geom_errorbarh(aes(xmin = Estimate-Est.Error, xmax = Estimate+Est.Error), height = 0, size = 5, alpha = 0.5) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), height = 0, size = 1) +
  geom_point() +
  labs(x = "Estimate", y = "Input", colour = "Input") +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

trials_trace <- trials %>% 
  as_tibble() %>% 
  select(group, input = word1, target = word2, phon_trace)

@@

# get top candidates
trace_top <- trace %>% 
  group_by(input, language, response) %>% 
  summarise(
    activation = max(activation, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  group_by(input, language) %>% 
  slice_max(activation, n = 10) %>% 
  ungroup() %>% 
  mutate(group = ifelse(language=="Catalan", "ENG-CAT", "ENG-SPA")) %>% 
  select(-language) %>% 
  # add experimental trials
  left_join(trials_trace) %>% 
  mutate(correct = response==phon_trace) %>% 
  select(group, input, response, activation, correct)

@@

ggplot(d_jtrace, aes(response_reordered, activation, fill = correct)) +
  facet_wrap(input_label~source, scales = "free_y", ncol = 2) +
  geom_col(colour = "white") +
  labs(x = "Top 10 candidates", y = "Max. TRACE activation") +
  coord_flip() +
  scale_x_reordered() +
  # scale_y_continuous(limits = c(0, 1)) +
  theme_bw() +
  theme(
    legend.position = "none"
  )

>>>>

---
title: "Translation Elicitation"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: 
---

@@

# import data
stimuli <- readRDS(here("Data", "stimuli.rds"))
participants <- readRDS(here("Data", "participants.rds"))
responses <- readRDS(here("Data", "responses.rds"))
jtrace <- readRDS(here("Data", "jtrace.rds"))

@@

```{r participants_stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

Participants in the ENG-SPA listened to `r nrow(stimuli[stimuli$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials. Five trials in each condition were practice trials and were excluded from analyses.

```{r stimuli_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
stimuli_coded <- stimuli %>% 
  mutate(
    test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"),
    zipf = log10(frequency)+3,
    pthn_log = log(pthn),
    lv_log = log(lv)
  )
```

@@

```{r stimuli_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  ggplot(aes(zipf, fill = test_language)) +
  geom_histogram(bins = 25) +
  labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme_custom() +
  theme(
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

@@

```{r stimuli_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  count(test_language, pthn) %>% 
  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
  geom_col(position = position_dodge()) +
  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

@@

```{r stimuli_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  ggplot(aes(group, lv, fill = group, colour = group)) +
  geom_violin() +
  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
  labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  ) +

  stimuli_coded %>% 
  select(trial_id, group, ends_with("_ratio")) %>% 
  pivot_longer(ends_with("_ratio"), names_to = "type", values_to = "ratio") %>% 
  mutate(type = str_to_sentence(str_remove(type, "_ratio"))) %>% 
  ggplot(aes(group, ratio, fill = group, colour = group)) +
  facet_wrap(~type, ncol = 1) +
  geom_violin() +
  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
  labs(x = "Grouop", y = "Similarity", fill = "Test language") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  ) +

@@

```{r stimuli_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  select(trial_id, group, onset) %>% 
  ggplot(aes(group, fill = onset, colour = onset)) +
  geom_bar(position = position_dodge(width = 0.9)) +
  labs(x = "Group", y = "# trials", fill = "Phonological onset", colour = "Phonological onset") +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

@@

```{r stimuli_similarity_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  select(trial_id, group, lv, onset) %>% 
  ggplot(aes(lv, fill = onset, colour = onset)) +
  facet_wrap(~group, ncol = 2) +
  geom_histogram() +
  labs(x = "Phonological Leveshtein distance\n(corrected for string length)", y = "# trials",
       fill = "Phonological onset", colour = "Phonological onset") +
  theme_custom() +
  theme(
    legend.position = c(0.75, 0.25),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

@@

m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
  mutate_at(vars(lv), function(x) paste0(x, " SD"))

ggplot(m, aes(pthn, .value, colour = lv, fill = lv)) +
  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(x = "Phon. neigh. dens. (SD)", y = "P(Correct)",
       colour = "Phon. similarity", fill = "Phon. similarity") +
  scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid = element_line(colour = "grey", linetype = "dotted")
  ) 
ggsave(here("Figures", "responses_pthn.png"))

@@

m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
  mutate_at(vars(pthn), function(x) paste0(x, " SD"))

ggplot(m, aes(lv, .value, colour = pthn, fill = pthn)) +
  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(x = "Phon. similarity (SD)", y = "P(Correct)", colour = "Phon. neigh. dens.",
       fill = "Phon. neigh. dens.") +
  scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    panel.grid = element_line(colour = "grey", linetype = "dotted")
  ) +
  ggsave(here("Figures", "responses_lv.png"), height = 4, width = 4.5)

@@

```{r jtrace_fixed_table, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
summary(fit_jtrace$fit)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_number(2:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "lv", "pthn:lv") | str_detect(term, "activation"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  )
```

```{r jtrace_fixed_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%"}
fit_jtrace$fixed %>%
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
  ggplot(aes(.value, fill = .variable)) +
  facet_wrap(~.variable, scales = "free") +
  geom_vline(xintercept = 0, colour = "black") +
  stat_halfeye() +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
    axis.text = element_text(size = 9)
  )
```

@@

stimuli_trace <- stimuli %>% 
  as_tibble() %>% 
  select(group, input = word1, target = word2, phon_trace)

@@

# get top candidates
trace_top <- trace %>% 
  group_by(input, language, response) %>% 
  summarise(
    activation = max(activation, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  group_by(input, language) %>% 
  slice_max(activation, n = 10) %>% 
  ungroup() %>% 
  mutate(group = ifelse(language=="Catalan", "ENG-CAT", "ENG-SPA")) %>% 
  select(-language) %>% 
  # add experimental trials
  left_join(stimuli_trace) %>% 
  mutate(correct = response==phon_trace) %>% 
  select(group, input, response, activation, correct)

@@

ggplot(d_jtrace, aes(response_reordered, activation, fill = correct)) +
  facet_wrap(input_label~source, scales = "free_y", ncol = 3) +
  geom_col(colour = "white") +
  labs(x = "Top 10 candidates", y = "Max. TRACE activation") +
  coord_flip() +
  scale_x_reordered() +
  # scale_y_continuous(limits = c(0, 1)) +
  theme_bw() +
  theme(
    legend.position = "none"
  )

### 2021-07-01 Rmd/te_manuscript.Rmd
Rename files, include job, add index.html

---
title             : "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
shorttitle        : "Phoneme-weighted Levenshtein Distance"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Second Author"
    affiliation   : "1,2"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Many studies in bilingualism have explored how form similarity between translation equivalents (i.e. cognate- ness) impacts language processing. Cognates range from being identical to sharing only a few key features. Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity. In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value. A widely used metric for word similarity is Levenshtein distance [@levenshtein1966binary]. This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other. However, a challenge when defining phonological similarity is that some phonemes (e.g. [k] and [g]) are perceptually more similar than others (e.g. [k] and [o]). The standard Lev- enshtein distance calculation treats each change equally. Thus it often overestimates phonological distance between words. We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g. vowel or consonant insertion) and position in the word. To estimate the weights, we are col- lecting behavioural data in an auditory translation elicitation task. Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words. The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent. Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation. The higher the probability, the higher the perceptual similarity. Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.

@@

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

@@

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

participants <- readRDS(here("Results", "participants.rds"))
stimuli <- readRDS(here("Results", "stimuli.rds"))
accuracy <- readRDS(here("Results", "accuracy.rds"))
fit <- readRDS(here("Results", "fit3_phon.rds"))
comparison <- readRDS(here("Results", "model_comparison.rds"))

@@

The experiment was implemented online using Psychopy/Pavlovia [@peirce2019]. Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla). First, participants were informed about the aims of the study and gave informed consent for participating. Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study. Third, participants completed the experimental task. Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants). They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible. English participants were randomly assigned to the list of Catalan or Spanish trials. Participants in the Catalan list were presented with `r count(stimuli, language)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, language)$n[2]` trials.

Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio ended. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word. Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish). The task took approximately 15 minutes to be completed.

@@

We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT). We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group. We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance. We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output. Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020]. Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013]. To test and account for cross-group differences, we included a random intercept for each group. We compared models using Paretho-Smoothed Importance Sampling (PSIS-LOO, @vehtari2016). More information about the models and model comparison can be found in Appendix X.

All analyses were performed in R environment [@rcoreteam2019]. We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016]. Audio duration, intensity and pitch were extracted using the `PraatR` R package [@albin2014].

>>>>

---
title             : "The role of cognateness in non-native spoken word recognition"
shorttitle        : "Cognateness and non-native word recognition"

author: 
  - name          : "Gonzalo Garcia-Castro"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Ramon Trial Fargas, 25-27, 08005 Barcelona"
    email         : "gonzalo.garciadecastro@upf.edu"
  - name          : "Serene Siow"
    affiliation   : "2"
  - name          : "Nuria Sebastian-Galles"
    affiliation   : "1"
  - name          : "Kim Plunkett"
    affiliation   : "2"

authornote: |
  GG and SS contributed equally and share first authorship.

abstract: |
  There is compelling evidence that bilinguals access their lexicon in a language non-selective way. For example, bilinguals produce and translated cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006). Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance. These measures partially ignore potential sources of variability during lexical access like vowel vs. consonant overlap, overlap in stressed syllables, onset overlap or the distance in features between both translations. In this study, we explored the impact of some of these variables on non-native word translation task: Spanish and English participants listened to non-native words (Catalan or Spanish) and were prompted to type their translation in their native language. Critically, participants where unfamiliar with the testing language, ensuring that they were not able to translate words based on previous knowledge on their meaning, leaving phonological information as the only cue participants were able to exploit to translate words to their native language correctly. We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.

@@

keywords          : "cognate, word recognition, translation, non-native, spoken word recognition"
wordcount         : "X"

bibliography      : ["r-references.bib", "references.bib"]

@@

```{r analysis-preferences, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

participants <- readRDS(here("Data", "participants.rds"))
stimuli <- readRDS(here("Data", "stimuli.rds"))
responses <- readRDS(here("Data", "responses.rds"))

@@

The experiment was implemented online using Psychopy/Pavlovia [@peirce2019]. Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla). First, participants were informed about the aims of the study and gave informed consent for participating. Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study. Third, participants completed the experimental task. Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants). They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible. English participants were randomly assigned to the list of Catalan or Spanish trials. Participants in the Catalan list were presented with `r count(stimuli, group)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, group)$n[2]` trials.

Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio ended. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word. Participants contributed a total of `r distinct(responses, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(responses, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(responses, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish). The task took approximately 15 minutes to be completed.

@@

We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT). We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group. We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance. We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output. Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020]. Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013]. To test and account for cross-group differences, we included a random intercept for each group. We compared models using leave-one-out cross-validation (LOO, @vehtari2016). More information about the models and model comparison can be found in Appendix X.

All analyses were performed in R environment [@rcoreteam2019]. We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
